---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bootgcompr

<!-- badges: start -->
<!-- badges: end -->

The `bootgcompr` package is a group of functions used in combination with the `gtsummary` package for calculating non-parametric bootstrapped statistics using the `boot` package. `bootgcompr` allows for calculating the following in direct combination with `gtsummary`:

- Mean and median summary statistics through `tbl_custom_summary()` 
- Mean difference (MD) by bootstrapped linear regression (both univariate and multivariate) directly in `tbl_summary()`
- Risk Ratio / Relative Risk (RR) by bootstrapped logistic regression directly in tbl_summary() using the `add_stat()` function.
- Absolute Risk Ratio / Absolute Risk Reduction (ARR/ARD) using G-computation directly in tbl_summary() using the `add_stat()` function.
- Associated 95% confidence intervals (95%CI) for all of the above mentioned statistics and analyses

## Installation

You can install the development version of bootgcompr from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mdcodedoctor/bootgcompr")
```

## Example of summary statistics

If necessary to calculate summary statistics, while ensuring that the outputted means (using the `mean_boot()` function) or medians (using the `median_boot()` function) are bootstrapped with their 95%CI the functions can be used with `tbl_custom_summary()` through the `stat_fns =` argument.

The function is called and specifies the number of applied resamples through the `R =` argument. If the number of resamples are not specified it uses the default number of 10,000 resamples.

```{r tbl_summary() example}
# load libraries
library(bootgcompr)
library(gtsummary)

# create tbl_custom_summary table with bootstrapped summary statistics
# set the `R = ` value to the number of wanted resamples. In this example we use 1000 resamples.
trial |> 
  tbl_custom_summary(
    include = c(ttdeath, marker),
    by = trt,
    statistic = ~ "{mean} ({conf.low}, {conf.high})",
    stat_fns = everything() ~ mean_boot(),
    missing = "no",
    type = everything() ~ "continuous"
    ) |> 
    add_overall()
```


