---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
```

# bootgcompr 0.1.1 (UNDER DEVELOPMENT)

<!-- badges: start -->

<!-- badges: end -->

The `bootgcompr` package is a group of functions used in combination
with the `gtsummary` package for calculating non-parametric bootstrapped
statistics using the `boot` package. `bootgcompr` allows for calculating
the following in direct combination with `gtsummary`:

**Currently implemented:**

-   Bootstrapped mean and median summary statistics through `tbl_custom_summary()`
    and the `stat_fns =` argument

-   Risk Ratio / Relative Risk (RR) by bootstrapped logistic regression
    directly in `tbl_summary()` using the `add_stat()` function.

-   Mean difference by bootstrapped linear regression or median
    difference using bootstrapped quantile regression directly through
    `tbl_summary()` using the `add_stat()` function.

-   Causal inference through G-computation for calculating risk
    difference, risk ratio, mean difference and ratio of means directly
    with `tbl_summary()` using the `add_stat()` function.
    Handles both binary/dichotomous and continuous outcomes, with or
    without adjustment for covariates.

-   95%CI and p-values for all of the above mentioned analyses. Standard
    errors and specific methods for each variable can easily be
    extracted from each table if necessary through the usual `gtsummary`
    functions.

-   Wrapper functions for easily formatting of tables including standard
    displaying of the underlying statistical method, including number of
    resamples used in each analysis and if the analyses were adjusted
    for covariates or not, including alphabetically displayed of which
    covariates were used in the model.

## NOTE

The package is currently under development and many functions might not
work as expected and may have errors. Please submit an issue to the
github repo, and I will address the issue if I can and when I have time.

The package is developed alongside current work obligations, and I
probably cannot provide a sufficient ETA. I am also open for any
requests and suggestions for the package, and will try to implement any
changes that could be relevant and if I have the skills for doing this. 

I am not a biostatistician nor a software engineer, so using the package
will be by your own responsibility. The package has had some parts of the code revied by a biostatistician with experience in `R`. Please evaluate the source code to
ensure what type of statistical test is actually applied to any analyses
you conduct if in doubt. 

## Installation

You can install the development version of bootgcompr from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mdcodedoctor/bootgcompr")
```

## Example of summary statistics

If necessary to calculate summary statistics, while ensuring that the
outputted means (using the `mean_boot()` function) or medians (using the
`median_boot()` function) are bootstrapped with their 95%CI the
functions can be used with `tbl_custom_summary()` through the
`stat_fns =` argument.

```{r table_summary, include = TRUE}
# load libraries
library(bootgcompr)
library(gtsummary)

# create table for bootstrapped mean summary statistics
table_summary <- trial |> 
  tbl_custom_summary(
    include = c(ttdeath, marker),
    by = trt,
    statistic = ~ "{mean} ({conf.low}, {conf.high})",
    stat_fns = everything() ~ mean_boot(R = 1000),
    missing = "no",
    type = everything() ~ "continuous"
    ) |> 
    add_overall()
```

```{r table_summary_print, include = TRUE}

# note: Neccessary to use as_flex_table() for the output in the .Rmd file. Works fine elsewhere with usual gtsummary output.

table_summary |> 
  gtsummary::as_flex_table()
```

## Example of G-computation

Estimation of absolute risk difference using G-computation as causal
inference, with adjustment for specified covariates using
`tbl_summary()` and the `add_gcomputation()` function.

```{r table_gcomputation, include = TRUE}
# create table for estimation of causal inference through G-computation
table_gcomputation <- trial |>
  tbl_summary(
    include = c(death),
    by = trt,
    statistic = everything() ~ "{p}%",
    missing = "no",
    type = everything() ~ "dichotomous",
  ) |>
  add_overall() |>
  add_ci(
    pattern = "{stat} ({ci})") |>
  add_gcomputation(adj.vars = c("age", "stage", "response"), 
                   R = 1000, 
                   ci_type = "bca", 
                   t.effect = "average", 
                   scale = "ratio", 
                   percentage = TRUE, 
                   footnotes_separate = TRUE)
```

```{r table_gcomputation_print, include = TRUE}

# note: Neccessary to use as_flex_table() for the output in the .Rmd file. Works fine elsewhere with usual gtsummary output.

table_gcomputation |> 
  gtsummary::as_flex_table()
```

## Acknowledgements

A huge kudos to the developers of all the packages used as dependencies in this package. Especially to the authors of `gtsummary`, a package which allows easy creation of summary and statistical tables for publications. Without the work they have completed, this package would not have existed.

## References

**NOTE: References are not fully updated, but will be addressed when
possible for all relevant packages and relevant publications**

Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J. Reproducible
summary tables with the gtsummary package. The R Journal 2021;13:570–80.
<https://doi.org/10.32614/RJ-2021-053>.

Angelo Canty, B. D. Ripley (2024). boot: Bootstrap R (S-Plus) Functions.
R package version 1.3-31.

A. C. Davison, D. V. Hinkley (1997). Bootstrap Methods and Their
Applications. Cambridge University Press, Cambridge. ISBN 0-521-57391-2,
<doi:10.1017/CBO9780511802843>.
