---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bootgcompr

<!-- badges: start -->

<!-- badges: end -->

The `bootgcompr` package is a group of functions used in combination
with the `gtsummary` package for calculating non-parametric bootstrapped
statistics using the `boot` package. `bootgcompr` allows for calculating
the following in direct combination with `gtsummary`:

**Currently implemented:**

-   Mean summary statistics through `tbl_custom_summary()`

-   Risk Ratio / Relative Risk (RR) by bootstrapped logistic regression
    directly in `tbl_summary()` using the `add_stat()` function

-   Absolute Risk Reduction / Absolute Risk Difference (ARR/ARD) using
    G-computation directly in `tbl_summary()` using the `add_stat()`
    function

-   Associated 95% confidence intervals (95%CI) for all of the above
    mentioned statistics and analyses

**Under development or yet to be implemented:**

-   Median summary statistics through `tbl_custom_summary()`

-   Mean difference (MD) by bootstrapped linear regression (both
    univariate and multivariate) directly in `tbl_summary()`

-   Associated 95% confidence intervals (95%CI) for all of the above
    mentioned statistics and analyses

## Installation

You can install the development version of bootgcompr from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mdcodedoctor/bootgcompr")
```

## Example of summary statistics

If necessary to calculate summary statistics, while ensuring that the
outputted means (using the `mean_boot()` function) or medians (using the
`median_boot()` function) are bootstrapped with their 95%CI the
functions can be used with `tbl_custom_summary()` through the
`stat_fns =` argument.

The function is called and specifies the number of applied resamples
through the `R =` argument. If the number of resamples are not specified
it uses the default number of 10,000 resamples.

```{r tbl_summary example, include = TRUE}
# load libraries
library(bootgcompr)
library(gtsummary)

# create table for bootstrapped mean summary statistics
trial |> 
  tbl_custom_summary(
    include = c(ttdeath, marker),
    by = trt,
    statistic = ~ "{mean} ({conf.low}, {conf.high})",
    stat_fns = everything() ~ mean_boot(R = 1000),
    missing = "no",
    type = everything() ~ "continuous"
    ) |> 
    add_overall() |> 
    as_gt() # note: as_gt() is called due to rendering issues in this document. This call is not necessary when using this                package.
```

## Acknowledgements

A huge kudos to the developers of `gtsummary`. Without their substantial
work, this package would not have existed.

## References

Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J. Reproducible
summary tables with the gtsummary package. The R Journal 2021;13:570–80.
<https://doi.org/10.32614/RJ-2021-053>.

Angelo Canty, B. D. Ripley (2024). boot: Bootstrap R (S-Plus) Functions.
R package version 1.3-31.

A. C. Davison, D. V. Hinkley (1997). Bootstrap Methods and Their
Applications. Cambridge University Press, Cambridge. ISBN 0-521-57391-2,
<doi:10.1017/CBO9780511802843>.
