---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
```

# bootgcompr

<!-- badges: start -->

<!-- badges: end -->

The `bootgcompr` package is a group of functions used in combination
with the `gtsummary` package for calculating non-parametric bootstrapped
statistics using the `boot` package. `bootgcompr` allows for calculating
the following in direct combination with `gtsummary`:

**Currently implemented:**

-   Mean and median summary statistics through `tbl_custom_summary()`
    and the `stat_fns =` argument

-   Risk Ratio / Relative Risk (RR) by bootstrapped logistic regression
    directly in `tbl_summary()` using the `add_stat()` function

-   Absolute Risk Reduction / Absolute Risk Difference (ARR/ARD) using
    G-computation directly in `tbl_summary()` using the `add_stat()`
    function.

-   Mean difference by bootstrapped linear regression or median
    difference using bootstrapped quantile regression directly through
    `tbl_summary()` using the `add_stat()` function.

-   Associated bootstrapped 95% confidence intervals (95%CI) and
    p-values for all of the above mentioned statistics and analyses

-   Wrapper functions for easily formatting of tables including standard
    displaying of the underlying statistical method, including number of
    resamples used in each analysis and if the analyses were adjusted
    for covariates or not, including alphabetically displayed of which
    covariates were used in the model.
    
    
## NOTE
The package is currently under development and many functions might not work as expected and may have errors. Please submit an issue to the github repo, and I will address the issue if I can and when I have time.

The package is developed alongside current work obligations, and I probably cannot provide a sufficient ETA. I am also open for any requests and suggestions for the package, and will try to implement any changes that could be relevant.

I am not a biostatistician nor a software engineer, so using the package will be by your own responsibility. Please evaluate the source code to ensure what type of statistical test is actually applied to any analyses you conduct.

## Installation

You can install the development version of bootgcompr from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mdcodedoctor/bootgcompr")
```

## Example of summary statistics

If necessary to calculate summary statistics, while ensuring that the
outputted means (using the `mean_boot()` function) or medians (using the
`median_boot()` function) are bootstrapped with their 95%CI the
functions can be used with `tbl_custom_summary()` through the
`stat_fns =` argument.

```{r table_summary, include = TRUE}
# load libraries
library(bootgcompr)
library(gtsummary)

# create table for bootstrapped mean summary statistics
table_summary <- trial |> 
  tbl_custom_summary(
    include = c(ttdeath, marker),
    by = trt,
    statistic = ~ "{mean} ({conf.low}, {conf.high})",
    stat_fns = everything() ~ mean_boot(R = 1000),
    missing = "no",
    type = everything() ~ "continuous"
    ) |> 
    add_overall()
```

```{r table_summary_print, include = TRUE}

# note: Neccessary to use as_flex_table() for the output in the .Rmd file. Works fine elsewhere with usual gtsummary output.

table_summary |> 
  gtsummary::as_flex_table()
```

## Example of G-computation

Estimation of absolute risk difference using G-computation as causal inference, with adjustment for specified covariates using `tbl_summary()` and the `add_gcomputation()` function.


```{r table_gcomputation, include = TRUE}
# create table for estimation of causal inference through G-computation
table_gcomputation <- trial |>
  tbl_summary(
    include = c(death),
    by = trt,
    statistic = everything() ~ "{p}%",
    missing = "no",
    type = everything() ~ "dichotomous",
  ) |>
  add_overall() |>
  add_ci(
    pattern = "{stat} ({ci})") |>
  add_gcomputation(adj.vars = c("stage", "age", "marker"), R = 1000, ci_type = "bca", percentage = TRUE)
```


```{r table_gcomputation_print, include = TRUE}

# note: Neccessary to use as_flex_table() for the output in the .Rmd file. Works fine elsewhere with usual gtsummary output.

table_gcomputation |> 
  gtsummary::as_flex_table()
```

## Acknowledgements

A huge kudos to the developers of all the packages used as dependencies in this package. Without the work they have completed, this package would not have existed. 

## References

**NOTE: References are not fully updated, but will be addressed when possible for all relevant packages and relevant publications**

Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J. Reproducible
summary tables with the gtsummary package. The R Journal 2021;13:570–80.
<https://doi.org/10.32614/RJ-2021-053>.

Angelo Canty, B. D. Ripley (2024). boot: Bootstrap R (S-Plus) Functions.
R package version 1.3-31.

A. C. Davison, D. V. Hinkley (1997). Bootstrap Methods and Their
Applications. Cambridge University Press, Cambridge. ISBN 0-521-57391-2,
<doi:10.1017/CBO9780511802843>.
